
---
title: 'w02'
author: 'Yekta Amirkhalili'
date: 'May 2nd, 2023'
output: pdf_document 
---


### Week 2 (Session 1) - May 15, 2023
**material to cover:**

---
**0.**
...**Data**...


```{r, data}

library(rlang)
library(dplyr)


data <- read.csv('movies.csv')


subData <- select(data, title, year)

head(subData, 5)
```


**1.**
...**Measure of location, (mean, median and mode)**...

***MEAN*** 
```{r, locations}

print('Want to know What is the average rating for movies made in a particular year?')
msg = ' Enter the Year: '

yr <- readline(prompt = msg)
yr <- as.integer(yr)

calc_mean <- function(year_){
    particular_year <- subset(data$imbd_rating, data$year == year_)
    
    rt_avg <- mean(particular_year)
    out <- list(particular_year, rt_avg)
    
    return(out)
} 

func_out <- calc_mean(yr)
output_list <- func_out[1]
output_avg <- func_out[2]

print('Here are some of the ratings from this year: ')
output_list

yr_str <- as.character(yr)

sprintf('The average ratings of the movies on this list from the year %s is %f .', yr_str, output_avg)

```
***MEDIAN***
```{r}
#same function, except calculate median 
calc_median <- function(year_){
    particular_year <- subset(data$imbd_rating, data$year == year_)
    
    rt_med <- median(particular_year)

    
    return(rt_med)
} 

median_out <- calc_median(yr)

sprintf('The Median ratings for year %s is %f .', yr_str, median_out)

```

***MODE***

```{r}
#same function, except calculate mode 
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

calc_mode <- function(year_){
    particular_year <- subset(data$imbd_rating, data$year == year_)
    
    rt_mod <- getmode(particular_year)

    
    return(rt_mod)
} 

mode_out <- calc_mode(yr)

sprintf('The Mode ratings for year %s is %f .', yr_str, mode_out)

```


...**Measure of variability, (variance, standard deviation)**...

```{r, variability}

calc_var <- function(year_){
    particular_year <- subset(data$imbd_rating, data$year == year_)
    
    rt_var <- var(particular_year)
    
    
    return(rt_var)
}


sprintf('The Variance of ratings for year %s is %f .', yr_str, calc_var(yr)) 

```

**2.**
~~~**Bayes Rule**~~~

*Introduction*

1. Bayes First Rule:

$$P(E) = \sum_{i = 1}^n P(E|F_{i})P(F_{i}) $$

> Ex1. Suppose we have two baskets. Basket A has 3 Red and 2 White balls in it. Basket B has 3 Red and 4 White balls in it. We randomly pick one ball from Basket A and transfer it to Basket B. Then we pick a ball, randomly, out of Basket B. What is the probability of the chosen ball being white? 



```{r}

bas_A = c('r','r','r', 'w','w')
bas_B = c('r','r','r', 'w','w','w','w')

pick_white_prob <- function(basket){
    #probability of picking white 
    total <- length(basket)
    basket_counts <- table(basket)
    whites <- basket_counts[names(basket_counts) == 'w']
    
    prob <- whites/total
    return(prob)
    
}

pick_red_prob <- function(basket){
    #probability of picking white 
    total <- length(basket)
    basket_counts <- table(basket)
    reds <- basket_counts[names(basket_counts) == 'r']
    
    prob <- reds/total
    return(prob)
    
}

```

* Scenario 1 
We pick 1 white ball from Basket A, transfer it to Basket B. Therefore, we now have 3 Red and 5 Whites in Basket B. 

* Scenario 2 
We pick 1 red ball from Basket A, transfer it to Basket B. Therefore, we now have 4 Red and 4 Whites in Basket B. 

In any case, the final probability calculation depends on what happened in the first pick. Let's see this in action! 

```{r, scn}

scenario <- function(num){
    #scenario 1 
    if(num == 1){
        new_b <- c('r','r','r','w','w','w','w', 'w')
    }else{ #scenario 2
        new_b <- c('r','r','r','r','w','w','w','w')
    }
    
    return(new_b)
    
}


```


```{r, probability}

newBasket1 <- scenario(1)
newBasket2 <- scenario(2)


prob_ <- function(a, b1, b2){
    
    # pick white from A, pick white from B of Scenario 1  
    total_prob_p1 <- pick_white_prob(a) * pick_white_prob(b1)
    
    #pick red from A, pick white from new B 
    total_prob_p2 <- pick_red_prob(a) * pick_white_prob(b2)
    
    probability <- total_prob_p1 + total_prob_p2
    
    return(probability)
    
}

prob_(bas_A, newBasket1, newBasket2)

```

**add**
...**Histograms**...

> Let's see the distribution of films in the Top 250 based on year of release. 

```{r}

release_date <- data$year 
hist(release_date)

```

```{r}

hist(release_date,
main="Year of Release for IMDB TOP 250 Films",
xlab="Release Date (year)",
col="#bfb12e",
freq=FALSE)

```

Save your Histogram in a file: 

```{r}
# Give the chart file a name.
png(file = "session1_histogram.png")

hist(release_date,
main="Year of Release for IMDB TOP 250 Films",
xlab="Release Date (year)",
col="#bfb12e")

# Save the file.
dev.off()
```







